Undirected Graph Exploration with $\Theta(\log\log n)$ Pebbles

### 2016-05-10 presentation/main.tex
mark for a nearly complete set of slides

\begin{frame}
  \frametitle{Exploring graphs with $\log\log n$ pebbles}
  \begin{itemize}
    \item There is a $(c,0,c\cdot\log n)$ pebble machine which explores
      a graph $G$ with $n$ vertices
    \item Applying above simulation $\log \log n$ times yields
      $(c'^{\log\log n}\cdot c, c'\cdot\log\log n, c)$ pebble machine
      \begin{theorem}
        Any bounded-degree graph on at most $n$ vertices can be explored using
        $\mathcal{O}(\log\log n)$ pebbles and memory.
      \end{theorem}
  \end{itemize}
\end{frame}

\section{Limitations of pebble machines}
\begin{frame}
  \frametitle{Limitations of pebbles}
  \begin{itemize}
    \item Construct 3-regular graphs which cannot be traversed by an agent
      $\mathcal{A}$ with $p$ pebbles
    \item Such graphs are called $p$-barriers
      \begin{center}
        \resizebox{0.7\textwidth}{!}{\input{tikz/barrierint}}
      \end{center}
    \item If $\mathcal{A}$ starts in any vertex in $G$ it will never traverse
      the barrier from $v_{1}$ ($v_{2}$) to $v_{2}$ or $v_{3}$ using at most
      $p$ pebbles
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Locality gadget}
  \begin{itemize}
    \item Replace edge by a $(p-1)$-barrier
    \item Enforce \enquote{locality} of pebbles
    \item An agent cannot traverse any gadget without \emph{all} his pebbles
    \item All pebbles of the agent are either carried or placed within the
      adjacent gadgets
  \end{itemize}
  \begin{center}
    \uncover<2->{\resizebox{0.8\textwidth}{!}{\input{tikz/locgadget}}}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Barrier construction: Initial thoughts}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Consider a subset of pebbles of cardinality $p$
      \item Use the locality gadget as edges
      \item Since pebbles are local: enumerate configurations in macro vertex
        $C_{1},\dots,C_{\alpha}$
      \item Project $A$ to a agent without pebbles $B$
      \item $B$ has $\alpha$ states and traverses through macro vertices
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/projection}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Barrier construction: The real deal}
  \begin{itemize}
    \item $B$ has no pebbles and is therefore prone to a $0$-barrier
    \item Apply locality gadget to all edges
    \item This yields a $p$-barrier for the chosen subset of pebbles
    \item Iterate $p$-barriers for all possible subsets
  \end{itemize}
  \begin{center}
    \resizebox{0.8\textwidth}{!}{\input{tikz/iterbarrier}}
  \end{center}
\end{frame}

>>>>

\begin{frame}
  \frametitle{Exploring graphs with $\log\log n$ pebbles}
  \begin{itemize}
    \item There is a $(c,0,c\cdot\log n)$ pebble machine which explores
      a graph $G$ with $n$ vertices
    \item Applying above simulation $\log \log n$ times yields
      $(c'^{\log\log n}\cdot c, c'\cdot\log\log n, c)$ pebble machine exploring
      $G$
      \begin{theorem}
        Any bounded-degree graph on at most $n$ vertices can be explored using
        $\mathcal{O}(\log\log n)$ pebbles and memory.
      \end{theorem}
  \end{itemize}
\end{frame}

\section{Limitations of pebble machines}
\begin{frame}
  \frametitle{Limitations of pebbles}
  \begin{itemize}
    \item Construct 3-regular graphs which cannot be traversed by an agent
      $\mathcal{A}$ with $p$ pebbles
    \item Such graphs are called $p$-barriers
      \begin{center}
        \resizebox{0.6\textwidth}{!}{\input{tikz/barrierint}}
      \end{center}
    \item $\mathcal{A}$ cannot traverse a barrier from $v_{1}$ or $v_{2}$
      to $v_{3}$ or $v_{4}$ or vice versa using at most $p$ pebbles
    \item Inductive construction of barriers
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Locality gadget}
  \begin{itemize}
    \item Replace edge by a $(p-1)$-barrier
    \item Enforce \enquote{locality} of pebbles
    \item An agent cannot traverse any gadget without \emph{all} his pebbles
    \item All pebbles of the agent are either carried or placed within the
      adjacent gadgets
      \begin{center}
        \uncover<2->{\resizebox{0.8\textwidth}{!}{\input{tikz/locgadget}}}
      \end{center}
    \item<3-> Original vertices are called \emph{macro} vertices
  \end{itemize}
\end{frame}

\begin{frame}
  \begin{theorem}
    Given an $(r-1)$-barrier $B$ with $m$ vertices for an agent $\mathcal{A}$
    with $p\geq r$ pebbles, we can construct an $r$-barrier $B'$ with
    $\mathcal{O}(\binom{p}{r}\cdot m\cdot\alpha_{B}^{2})$ vertices for
    $\mathcal{A}$.
  \end{theorem}
  \begin{itemize}
    \item $\alpha_{B}$ is the amount of \enquote{local} configurations for
      $\mathcal{A}$ with $r$ pebbles
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Barrier construction: Initial thoughts}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Consider a subset of pebbles of cardinality $r$
      \item Use the locality gadget as edges
      \item Since pebbles are local: enumerate configurations in macro vertex
        $C_{1},\dots,C_{\alpha}$
      \item Project $A$ to a agent without pebbles $B$
      \item $B$ has $\alpha$ states and traverses through macro vertices
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/projection}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Barrier construction}
  \begin{itemize}
    \item $B$ has no pebbles and is therefore prone to a $0$-barrier
    \item Apply locality gadget to all edges
    \item This yields a $p$-barrier for the chosen subset of pebbles
    \item Iterate $p$-barriers for all possible subsets
      \begin{center}
        \resizebox{0.7\textwidth}{!}{\input{tikz/iterbarrier}}
      \end{center}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Barrier base case}
  \begin{theorem}
    For any $q$ non cooperative $s$-state agents without pebbles, there
    exists a 3-regular graph $G$ on $\mathcal{O}(qs)$ vertices with the
    following property: There are two edges $\left\{v_{1},v_{2}\right\}$
    and $\left\{v_{3},v_{4}\right\}$ in $G$, the first labeled 0, such
    that any of the $q$ agents starting in $v_{1}$ or $v_{2}$ does not
    traverse the edge $\left\{v_{3},v_{4}\right\}$.
  \end{theorem}
  \begin{itemize}
    \item For agent $\mathcal{A}$ consider
      $S = \left\{\mathcal{A}_{q}\mid q\in Q\right\}$
    \item Since $|S| = |Q| = q$ there is a 0-barrier with $\mathcal{O}(s^{2})$
      vertices
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Barrier}
  \begin{theorem}
    Given an $(r-1)$-barrier $B$ with $m$ vertices for an agent $\mathcal{A}$
    with $p\geq r$ pebbles, we can construct an $r$-barrier $B'$ with
    $\mathcal{O}(\binom{p}{r}\cdot m\cdot\alpha_{B}^{2})$ vertices for
    $\mathcal{A}$.
  \end{theorem}
  \begin{center}
    \resizebox{0.7\textwidth}{!}{\input{tikz/iterbarrier}}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Barrier size}
  \begin{itemize}
    \item Iterative construction
  \end{itemize}
  \begin{center}
    \resizebox{0.8\textwidth}{!}{\input{tikz/barriersize}}
  \end{center}
  \begin{theorem}
    For $r\leq p$ and $s\geq 2^{p}$, the number of vertices of $r$-barrier
    $B_{r}$ for the $s$-state agent $\mathcal{A}$ with $p$ pebbles is bounded
    by $\mathcal{O}(s^{8^{r+1}})$.
  \end{theorem}
\end{frame}

\begin{frame}
  \frametitle{Lower pebble bound}
  \begin{theorem}
    For any constant $\epsilon > 0$, an agent with at most
    $\mathcal{O}((\log n)^{1-\epsilon})$ bits of memory needs at least
    $\Omega(\log\log n)$ distinguishable pebbles for exploring all
    graphs on at most $n$ vertices.
  \end{theorem}
\end{frame}
